---
title: "Spatial Stream Network Analysis of Nechako Watershed Stream Temperatures 2022b"
author: "Hill, N.H., Thorley, J.L., & Irvine, A."
bibliography: bibliography.bib
---
  
```{r, echo = FALSE, warning = FALSE, message = FALSE, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, comment = NA, results = "asis", cache = FALSE)

source("header.R")
library(poisreport)

nheaders <- 2L

rename <- c(
  "map" = "Maps", # to avoid rendering problems!
  "temperature-air2stream" = "Stream Temperature"
)

sort <- c(
  "description", "coef", "glance", "map",
  "sensitivity", "sensitivity_parameter", "sensitivity_term",
  "covariance-distance", "air-water-temp", "water-temp",
  "gsdd-annual-site", "gsdd-map"
)

drop <- c("baseline", "temperature-mechanistic", "temperature")
```

```{r}
cat(report_draft())
```

The suggested citation for this [analytic appendix](https://www.poissonconsulting.ca/analytic-appendices.html) is:
  
```{r}
cat(report_citation(file_name()))
```

## Background

The primary goal of the current analyses is to answer the following question:

  > How can we model stream temperature to include spatial correlation through a stream network?
  
### Data Preparation
  
Sub-hourly water temperature data collected in the Nechako Watershed in northern British Columbia between 2019 and 2021 were downloaded as csv files from [Zenodo](https://zenodo.org/records/6426024#.ZEAqr-zMI0Q) [@morris_sub-hourly_2022].

Hourly air temperature data (at two metres above ground level) for the Nechako Watershed for the years 2019-2021 were downloaded from the [ERA-5-Land simulation](https://cds.climate.copernicus.eu/cdsapp#!/dataset/10.24381/cds.e2161bac?tab=overview) using the Copernicus Climate Change Service (C3S) Climate Data Store as NetCDF files [@munoz_sabater_era5-land_2019].

Daily baseflow and surface runoff data for the Nechako Watershed for the years 2019-2021 using the ACCESS1-0_rcp85 climate scenario were downloaded from the [Pacific Climate Impacts Consortium's Gridded Hydrologic Model Output](https://www.pacificclimate.org/data/gridded-hydrologic-model-output) as NetCDF files [@pacific_climate_impacts_consortium_university_of_victoria_gridded_2020].

The data were prepared for analysis using `r substr(R.version.string,1,15)` [@r_core_team_r_2022].

Key assumptions of the data preparation included:
  
- All stream temperature data are correct, except those flagged "Fail", which were excluded from analysis (see @gilbert_sub-hourly_2022 for details).
- The simulated air temperature and discharge data from their respective modeled simulations are reasonable approximations of the truth.
- The daily discharge at each stream temperature site was calculated by taking the weighted average of the sum of the baseflow and runoff for the watershed area upstream of the site; these were then averaged over weekly periods for each site.
- One site that corresponded to just four consecutive observed data points proved insufficient to capture the annual-scale fluctuations in stream temperature; this site (WHC) was dropped from the analysis.
- There were 146 observations with negative stream temperatures measurements, which were all set to 0˚C.

### Statistical Analysis

Model parameters were estimated using Bayesian methods.
The estimates were produced using Stan [@carpenter_stan_2017].
For additional information on Bayesian estimation the reader is referred to @mcelreath_statistical_2020.

Unless stated otherwise, the Bayesian analyses used weakly informative prior distributions [@gelman_prior_2017].
The posterior distributions were estimated from `r getOption("mb.nchains", 3L) * getOption("mb.niters", 500L) ` Markov Chain Monte Carlo (MCMC) samples thinned from the second halves of `r getOption("mb.nchains", 3L)` chains [@kery_bayesian_2011, pp. 38-40].
Model convergence was confirmed by ensuring that the potential scale reduction factor $\hat{R} \leq `r round(getOption("mb.rhat", 1.05), 2)`$ [@kery_bayesian_2011, pp. 40] and the effective sample size [@brooks_handbook_2011] $\textrm{ESS} \geq  `r getOption("mb.nchains", 3L) * getOption("mb.niters", 500) * round(getOption("mb.esr", 0.1), 2)`$ for each of the monitored parameters [@kery_bayesian_2011, pp. 61].

The sensitivity of the posteriors to the choice of prior distributions was evaluated by doubling the standard deviations of the priors and then using $\hat{R}$ to evaluate whether the samples were drawn from the same posterior distribution [@thorley_fishing_2017].

The parameters are summarized in terms of the point *estimate*, *lower* and *upper* 95% compatibility limits [@rafi_semantic_2020] and the surprisal *s-value* [@greenland_valid_2019].
Together a pair of lower and upper compatibility limits (CLs) are referred to as a compatibility interval (CI).
The estimate is the median (50th percentile) of the MCMC samples while the 95% CLs are the 2.5th and 97.5th percentiles.
The s-value indicates how surprising it would be to discover that the true value of the parameter is in the opposite direction to the estimate [@greenland_valid_2019].
An s-value of $>$ 4.32 bits, which is equivalent to a p-value $<$ 0.05 [@kery_bayesian_2011; @greenland_living_2013], indicates that the surprise would be equivalent to throwing at least 4.3 heads in a row on a fair coin.

Variable selection was based on the heuristic of directional certainty [@kery_bayesian_2011; @murtaugh_defense_2014; @castilho_towards_2021].
Fixed effects were included if their s-value was $>$ 4.32 bits [@kery_bayesian_2011].
Based on a similar argument, random effects were included if their standard deviation had a lower 95% CL $>$ 5% of the median estimate.

The analyses were implemented using `r substr(R.version.string,1,15)` [@r_core_team_r_2022] and the [`embr`](https://github.com/poissonconsulting/embr) family of packages.

### Model Descriptions

#### Stream Temperature

The data were analyzed using a Spatial Stream Network model [@ver_hoef_moving_2010; @peterson_mixedmodel_2010], with code adapted from the [`SSNbayes`](https://github.com/EdgarSantos-Fernandez/SSNbayes) package [@santos-fernandez_bayesian_2022].
The necessary stream network distances and connectivity were calculated using the BC Freshwater Atlas. 
Air and stream temperature data were averaged by site and week; modeling was done on this weekly time scale.

The expected stream temperatures were modeled using the 4-parameter version of the air2stream model [@toffolon_hybrid_2015].
The average stream temperature (in ˚C) in the first week, $W_{s,j=1}$ was estimated by the model and assumed to be the same for all sites.

For all subsequent weeks (i.e., $j > 1$), the change in the stream temperature (in ˚C) between week $j - 1$ and week $j$ for the $s^{th}$ site, $\Delta W_{s,j}$, was modeled as follows:

\begin{equation}
\Delta W_{s,j} = \frac{1}{(\theta_{s,j})^{a4_s}}(a1_s + a2_s A_{s,j} - a3_s W_{s,j - 1}),
\end{equation}

where $a1_s$, $a2_s$, $a3_s$, and $a4_s$ are the parameters of the air2stream model for the $s^{th}$ site, $A_{s,j}$ is the air temperature (in ˚C) for the $s^{th}$ site in the $j^{th}$ week, $W_{s, j - 1}$ is the expected stream temperature at the $s^{th}$ site in the previous week, and $\theta_{s,j}$ is the dimensionless discharge for the $s^{th}$ site in the $j^{th}$ week.
$\theta_{s,j}$ was calculated as follows:

\begin{equation}
\theta_{s,j} = \frac{d_{s,j}}{\bar{d_s}}
\end{equation}

where $d_{s,j}$ is the discharge for the $s^{th}$ site in the $j^{th}$ week, and $\bar{d_s} = \frac{\sum_{j = 1}^{J}(d_{s,j})}{J}$ is the mean discharge across all $J$ weeks for the $s^{th}$ site.

The expected stream temperature for the $s^{th}$ site in the $j^{th}$ week was then calculated:

\begin{equation}
W_{s,j} = W_{s,j - 1} + \Delta W_{s,j}.
\end{equation}

Growing Season Degree Days (GSDD) are the accumulated thermal units (in ˚C) during the growing season based on the mean daily water temperature values, which is a useful predictor of age-0 rainbow and westslope cutthroat trout size at the beginning of winter.
The start and end of the growing season were based on the definitions of @coleman_cold_2007:

- Start: the beginning of the first week that average stream temperatures exceeded and remained above 5˚C for the season. 
- End: the last day of the first week that average stream temperature dropped below 4˚C.

GSDD were derived for each site and year by assuming that the daily stream temperatures at each site were the predicted weekly mean stream temperature for every day in the given week.

Key assumptions of the model include:

- The stream network is dendritic, not braided.
- The expected stream temperatures were set to 0˚C if they were estimated to be negative by the model.
- The stream temperature in the first week is the same for all sites.
- The parameters of the air2stream model ($a1$, $a2$, $a3$, and $a4$) vary randomly by site.
- The residual variation is multivariate normally distributed.
- The covariance structure of the residual variation combines the following covariance components:
  - Nugget (allows for variation at a single location) 
  - Exponential tail-down (allows for spatial dependence between flow-connected and flow-unconnected locations)

Preliminary analysis found that:

- The exponential tail-down model was better at explaining the spatial correlation in the data than exponential tail-up or euclidean distance models [@ver_hoef_moving_2010; @peterson_mixedmodel_2010].
- The full 8-parameter air2stream model did not converge.
- Preliminary analysis found that allowing the initial stream temperature to vary randomly by site produced unrealistic stream temperatures for January (> 10˚C).

### Model Templates

```{r}
cat(sbr_blocks(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

## Results

### Tables

```{r}
cat(sbr_tables(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

### Figures

```{r}
cat(sbr_figures(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

## Acknowledgements

The organisations and individuals whose contributions have made this analytic appendix possible include:
  
- Hillcrest Geographics
  - Simon Norris

\newpage

## References
